---
title: "Data Overview"
author: "XSun"
date: "2023-09-25"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=F, message=FALSE, warning=FALSE}
library(GenomicRanges)
library(ggplot2)
```

# Introduction

We compare some new ATAC-seq data sets with our previous data here. 


# MI

Spatial multi-omic map of human myocardial infarction

Kuppe, C., Ramirez Flores, R.O., Li, Z. et al. Spatial multi-omic map of human myocardial infarction. Nature 608, 766–777 (2022). https://doi.org/10.1038/s41586-022-05060-x

downloaded from 
https://cellxgene.cziscience.com/collections/8191c283-0816-424b-9b61-c3e1d6258a77

sub-cluster downloaded from https://zenodo.org/record/7098004#.Y0P_LC0RoeY

31 samples


|             | # of cell (*1000, MI) | # of cell (*1000, A&K) | # of peaks (MI, not cell type specific) | # of peaks (A&K, not cell type specific) | # of overlaps (not cell type specific) |
|-------------|-----------------------|------------------------|-----------------------------------------|------------------------------------------|----------------------------------------|
| Total       | 46.2                  | 26.09                  |                                         |                                          |                                        |
| vCMs        | 22                    | 9.4                    | 182,116                                 | 150,000                                  | 140,289                                |
| fibroblasts | 7                     | 5.4                    | 173,843                                 | 150,000                                  | 129,347                                |
| endothelial | 7                     | 5.4                    | 87,364                                  | 120,662                                  | 85,426                                 |
| myeloid     | 4.5                   | 1.6                    | 83,560                                  | 98,167                                   | 13,303                                 |
| pericytes   | 3.5                   | 3.0                    | 138,683                                 | 109,447                                  | 101,936                                |
| lymphoid    | 1.2                   | 0.5                    | 111,685                                 | 44,995                                   | 45,286                                 |
| vSMCs       | 0.6                   | 0.35                   | 94,685                                  | 44,997                                   | 46,683                                 |
| neuronal    | 0.4                   | 0.44                   | 91,725                                  | 55,483                                   | 49,272                                 |


# CHD

Integrated multi-omic characterization of congenital heart disease

https://www.nature.com/articles/s41586-022-04989-3#data-availability

Downloaded from https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE203273

Bulk data

42 samples

|      | # of peaks |
|------|------------|
| cd14 | 666,618    |
| nk   | 6,745,378  |
| pcm1 | 166,035    |

``` {r echo=F, message=FALSE, warning=FALSE,  fig.width=10}

load("/project2/gca/xsun/heart_proj/data/CHD/atac/processed/cd14_gr_hg19.rdata")
element_widths <- width(merged_bigwig_reduced)
# Assuming element_widths is the vector you obtained earlier
# Create a data frame for plotting
data <- data.frame(widths = element_widths)

# Create the histogram using ggplot2
ggplot(data, aes(x = widths)) +
  geom_histogram() +  # Adjust binwidth as needed
  labs(x = "Peak Width", y = "Count") +
  ggtitle("CD14")

# Calculate the 90th percentile
percentile_80 <- quantile(element_widths, 0.80)

# Keep only the elements below the 90th percentile
filtered_widths <- element_widths[element_widths < percentile_80]

# Check the length of the filtered vector
cat("Number of elements after removing the top 20%:", length(filtered_widths), "\n")

# Create a histogram of the filtered element widths using ggplot2
ggplot(data.frame(widths = filtered_widths), aes(x = widths)) +
  geom_histogram() +  # Adjust binwidth as needed
  labs(x = "Peak Width", y = "Count") +
  ggtitle("CD14 (Excluding Top 20%)")


load("/project2/gca/xsun/heart_proj/data/CHD/atac/processed/nk_gr_hg19.rdata")
element_widths <- width(merged_bigwig_reduced)
# Assuming element_widths is the vector you obtained earlier
# Create a data frame for plotting
data <- data.frame(widths = element_widths)

# Create the histogram using ggplot2
ggplot(data, aes(x = widths)) +
  geom_histogram() +  # Adjust binwidth as needed
  labs(x = "Peak Width", y = "Count") +
  ggtitle("NK")

# Calculate the 90th percentile
percentile_80 <- quantile(element_widths, 0.80)

# Keep only the elements below the 90th percentile
filtered_widths <- element_widths[element_widths < percentile_80]

# Check the length of the filtered vector
cat("Number of elements after removing the top 20%:", length(filtered_widths), "\n")

# Create a histogram of the filtered element widths using ggplot2
ggplot(data.frame(widths = filtered_widths), aes(x = widths)) +
  geom_histogram() +  # Adjust binwidth as needed
  labs(x = "Peak Width", y = "Count") +
  ggtitle("NK (Excluding Top 20%)")

load("/project2/gca/xsun/heart_proj/data/CHD/atac/processed/pcm1_gr.rdata")
element_widths <- width(merged_bigwig_reduced)
# Assuming element_widths is the vector you obtained earlier
# Create a data frame for plotting
data <- data.frame(widths = element_widths)

# Create the histogram using ggplot2
ggplot(data, aes(x = widths)) +
  geom_histogram() +  # Adjust binwidth as needed
  labs(x = "Peak Width", y = "Count") +
  ggtitle("PCM1")

# Calculate the 90th percentile
percentile_80 <- quantile(element_widths, 0.80)

# Keep only the elements below the 90th percentile
filtered_widths <- element_widths[element_widths < percentile_80]

# Check the length of the filtered vector
cat("Number of elements after removing the top 20%:", length(filtered_widths), "\n")

# Create a histogram of the filtered element widths using ggplot2
ggplot(data.frame(widths = filtered_widths), aes(x = widths)) +
  geom_histogram() +  # Adjust binwidth as needed
  labs(x = "Peak Width", y = "Count") +
  ggtitle("PCM1 (Excluding Top 20%)")

```


# Immune cells

Landscape of stimulation-responsive chromatin across diverse human immune cells

https://www.nature.com/articles/s41588-019-0505-9

8 healthy donors

collected by Jing Gu, storaged in /project2/xinhe/shared_data/AADCRC_lung/Caldero2019

## re-grouped by major cell types

| Cell types              | # of peaks | Cell types             | # of peaks | Cell types                   | # of peaks |
|-------------------------|------------|------------------------|------------|------------------------------|------------|
| Bcell_merge.bed         | 256235     | Bcell_rest.bed         | 206835     | Bcell_stimulated.bed         | 222088     |
| CD4_T_merge.bed         | 337358     | CD4_T_rest.bed         | 227545     | CD4_T_stimulated.bed         | 314472     |
| CD8_T_merge.bed         | 249173     | CD8_T_rest.bed         | 159600     | CD8_T_stimulated.bed         | 236933     |
| Gamma_delta_T_merge.bed | 151743     | Gamma_delta_T_rest.bed | 103283     | Gamma_delta_T_stimulated.bed | 135191     |
| Myeloid_merge.bed       | 301735     | Myeloid_rest.bed       | 276244     |                              |            |
| NK_merge.bed            | 243451     | NK_rest.bed            | 218348     | NK_stimulated.bed            | 180543     |
| T_cells_merge.bed       | 351149     |                        |            |                              |            |


## smaller groups

| rest                         |        | stimulated                   |        | differential peaks        |       |
|------------------------------|--------|------------------------------|--------|---------------------------|-------|
| Bcell_group.bed              | 206835 | Bcell_group.bed              | 222088 |                           |       |
| Bulk_B.bed                   | 127405 | Bulk_B.bed                   | 140859 |                           |       |
| CD4_T_group.bed              | 227545 | CD4_T_group.bed              | 314472 |                           |       |
| CD8_T_group.bed              | 159600 | CD8_T_group.bed              | 236933 |                           |       |
| CD8pos_T.bed                 | 134477 | CD8pos_T.bed                 | 135808 |                           |       |
| Central_memory_CD8pos_T.bed  | 95884  | Central_memory_CD8pos_T.bed  | 148500 | Central_memory_CD8pos_T_U | 924   |
| Effector_CD4pos_T.bed        | 128080 | Effector_CD4pos_T.bed        | 147309 | Effector_CD4pos_T_U       | 610   |
| Effector_memory_CD8pos_T.bed | 98532  | Effector_memory_CD8pos_T.bed | 140786 |                           |       |
| Follicular_T_Helper.bed      | 127264 | Follicular_T_Helper.bed      | 225183 |                           |       |
| Gamma_delta_T.bed            | 103283 | Gamma_delta_T.bed            | 135191 |                           |       |
| Immature_NK.bed              | 182210 |                              |        | Immature_NK_U             | 6832  |
| Mature_NK.bed                | 136860 | Mature_NK.bed                | 180543 |                           |       |
| Mem_B.bed                    | 137766 | Mem_B.bed                    | 189137 | Mem_B_U                   | 2803  |
| Memory_NK.bed                | 137181 |                              |        |                           |       |
| Memory_Teffs.bed             | 107882 | Memory_Teffs.bed             | 157724 | Memory_Teffs_U            | 200   |
| Memory_Tregs.bed             | 102609 | Memory_Tregs.bed             | 161057 |                           |       |
| Monocytes.bed                | 226420 | Monocytes.bed                | 211106 | Monocytes_U               | 38354 |
| Myeloid_DCs.bed              | 140685 |                              |        |                           |       |
| Myeloid_group.bed            | 276244 | Myeloid_group.bed            | 211106 |                           |       |
| NK_group.bed                 | 218348 | NK_group.bed                 | 180543 |                           |       |
| Naive_B.bed                  | 139387 | Naive_B.bed                  | 154570 | Naive_B_U                 | 423   |
| Naive_CD8_T.bed              | 96921  | Naive_CD8_T.bed              | 196974 | Naive_CD8_T_U             | 5029  |
| Naive_Teffs.bed              | 128112 | Naive_Teffs.bed              | 165885 | Naive_Teffs_U             | 2405  |
| Naive_Tregs.bed              | 137242 | Naive_Tregs.bed              | 174947 | Naive_Tregs_U             | 140   |
| Plasmablasts.bed             | 162676 |                              |        |                           |       |
| Regulatory_T.bed             | 121420 | Regulatory_T.bed             | 167853 |                           |       |
| Th17_precursors.bed          | 106904 | Th17_precursors.bed          | 158520 |                           |       |
| Th1_precursors.bed           | 110248 | Th1_precursors.bed           | 177126 |                           |       |
| Th2_precursors.bed           | 120143 | Th2_precursors.bed           | 173315 |                           |       |
| pDCs.bed                     | 143856 |                              |        | pDCs_U                    | 19308 |
