---
title: "Data Overview"
author: "XSun"
date: "2023-09-25"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=F, message=FALSE, warning=FALSE}
library(GenomicRanges)
library(ggplot2)
```

# Introduction

We compare some new ATAC-seq data sets with our previous data here. 


# MI

Spatial multi-omic map of human myocardial infarction

Kuppe, C., Ramirez Flores, R.O., Li, Z. et al. Spatial multi-omic map of human myocardial infarction. Nature 608, 766–777 (2022). https://doi.org/10.1038/s41586-022-05060-x

downloaded from 
https://cellxgene.cziscience.com/collections/8191c283-0816-424b-9b61-c3e1d6258a77

sub-cluster downloaded from https://zenodo.org/record/7098004#.Y0P_LC0RoeY


|             | # of cell (*1000, MI) | # of cell (*1000, A&K) | # of peaks (MI, not cell type specific) | # of peaks (A&K, not cell type specific) | # of overlaps (not cell type specific) |
|-------------|-----------------------|------------------------|-----------------------------------------|------------------------------------------|----------------------------------------|
| Total       | 46.2                  | 26.09                  |                                         |                                          |                                        |
| vCMs        | 22                    | 9.4                    | 182,116                                 | 150,000                                  | 140,289                                |
| fibroblasts | 7                     | 5.4                    | 173,843                                 | 150,000                                  | 129,347                                |
| endothelial | 7                     | 5.4                    | 87,364                                  | 120,662                                  | 85,426                                 |
| myeloid     | 4.5                   | 1.6                    | 83,560                                  | 98,167                                   | 13,303                                 |
| pericytes   | 3.5                   | 3.0                    | 138,683                                 | 109,447                                  | 101,936                                |
| lymphoid    | 1.2                   | 0.5                    | 111,685                                 | 44,995                                   | 45,286                                 |
| vSMCs       | 0.6                   | 0.35                   | 94,685                                  | 44,997                                   | 46,683                                 |
| neuronal    | 0.4                   | 0.44                   | 91,725                                  | 55,483                                   | 49,272                                 |


# CHD

Integrated multi-omic characterization of congenital heart disease

https://www.nature.com/articles/s41586-022-04989-3#data-availability

Downloaded from https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE203273

Bulk data

|      | # of peaks |
|------|------------|
| cd14 | 666,618    |
| nk   | 6,745,378  |
| pcm1 | 166,035    |

``` {r echo=F, message=FALSE, warning=FALSE,  fig.width=10}

load("/project2/gca/xsun/heart_proj/data/CHD/atac/processed/cd14_gr_hg19.rdata")
element_widths <- width(merged_bigwig_reduced)
# Assuming element_widths is the vector you obtained earlier
# Create a data frame for plotting
data <- data.frame(widths = element_widths)

# Create the histogram using ggplot2
ggplot(data, aes(x = widths)) +
  geom_histogram() +  # Adjust binwidth as needed
  labs(x = "Peak Width", y = "Count") +
  ggtitle("CD14")

# Calculate the 90th percentile
percentile_80 <- quantile(element_widths, 0.80)

# Keep only the elements below the 90th percentile
filtered_widths <- element_widths[element_widths < percentile_80]

# Check the length of the filtered vector
cat("Number of elements after removing the top 20%:", length(filtered_widths), "\n")

# Create a histogram of the filtered element widths using ggplot2
ggplot(data.frame(widths = filtered_widths), aes(x = widths)) +
  geom_histogram() +  # Adjust binwidth as needed
  labs(x = "Peak Width", y = "Count") +
  ggtitle("CD14 (Excluding Top 20%)")


load("/project2/gca/xsun/heart_proj/data/CHD/atac/processed/nk_gr_hg19.rdata")
element_widths <- width(merged_bigwig_reduced)
# Assuming element_widths is the vector you obtained earlier
# Create a data frame for plotting
data <- data.frame(widths = element_widths)

# Create the histogram using ggplot2
ggplot(data, aes(x = widths)) +
  geom_histogram() +  # Adjust binwidth as needed
  labs(x = "Peak Width", y = "Count") +
  ggtitle("NK")

# Calculate the 90th percentile
percentile_80 <- quantile(element_widths, 0.80)

# Keep only the elements below the 90th percentile
filtered_widths <- element_widths[element_widths < percentile_80]

# Check the length of the filtered vector
cat("Number of elements after removing the top 20%:", length(filtered_widths), "\n")

# Create a histogram of the filtered element widths using ggplot2
ggplot(data.frame(widths = filtered_widths), aes(x = widths)) +
  geom_histogram() +  # Adjust binwidth as needed
  labs(x = "Peak Width", y = "Count") +
  ggtitle("NK (Excluding Top 20%)")

load("/project2/gca/xsun/heart_proj/data/CHD/atac/processed/pcm1_gr.rdata")
element_widths <- width(merged_bigwig_reduced)
# Assuming element_widths is the vector you obtained earlier
# Create a data frame for plotting
data <- data.frame(widths = element_widths)

# Create the histogram using ggplot2
ggplot(data, aes(x = widths)) +
  geom_histogram() +  # Adjust binwidth as needed
  labs(x = "Peak Width", y = "Count") +
  ggtitle("PCM1")

# Calculate the 90th percentile
percentile_80 <- quantile(element_widths, 0.80)

# Keep only the elements below the 90th percentile
filtered_widths <- element_widths[element_widths < percentile_80]

# Check the length of the filtered vector
cat("Number of elements after removing the top 20%:", length(filtered_widths), "\n")

# Create a histogram of the filtered element widths using ggplot2
ggplot(data.frame(widths = filtered_widths), aes(x = widths)) +
  geom_histogram() +  # Adjust binwidth as needed
  labs(x = "Peak Width", y = "Count") +
  ggtitle("PCM1 (Excluding Top 20%)")

```
