---
title: "Differential and enrichment analysis"
author: "XSun"
date: "2024-05-28"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

# Introduction

## Differential Analysis

We did differential analysis for the topics from [k=10 and k=11](https://xsun1229.github.io/heart_project/fasttopic_running.html#Finalized_Results_–_cells_with_less_than_1000_peak_accessibility_removed,_counts_binarized). 

The differential analysis was performed under default settings using the `fastTopics` package.

```{r eval=F}
de <- de_analysis(fit, counts, pseudocount = 0.1,
                  control = list(ns = 1e4, nc = 10))
```

## Mapping ATAC-seq Peaks to Genes

We mapped the peaks to the closest protein-coding genes using `EnsDb.Hsapiens.v86`.

Peaks that are 500kb away from a gene were excluded in the enrichment analysis.

```{r eval=F}
closest_genes <- ClosestFeature(fib_atac, regions = granges(fib_atac))

```

```{r}
load("/project/xinhe/xsun/heart_project/data/fib_nearest_genes_proteincoding.rdata")
head(closest_genes)
```

## Pathway Enrichment Analysis

The enrichment analysis was performed using the `WebGestaltR` package with the following function:

```{r echo=TRUE, eval=FALSE}
enrich_Res <- WebGestaltR(enrichMethod="ORA", organism="hsapiens",
                          interestGene=genes_pass,
                          referenceGene=baseline_genes,
                          enrichDatabase="geneontology_Biological_Process",
                          interestGeneType="genesymbol",
                          referenceGeneType="genesymbol", isOutput=F)
```                          

Here, genes_pass are the genes passing a certain threshold, and baseline_genes are the genes included in our dataset.

# Results

## k=10

### Differential analysis

The topic 1 is likely to be the unique topic for myofib

```{r}
load("/project/xinhe/xsun/heart_project/1.fast_topic/results/fast_de_modified_k10.0cellremoved.1krm.rdata")
load("/project/xinhe/xsun/heart_project/1.fast_topic/plots/volcano_plot_k10_topic1.rdata")
print(volcano_plot)
```

### Pathway enrichment analysis

We select genes of interests using a relaxed threshold: 

lfsr_tp < 0.05 & z_tp > 5, lfc_tp >1

The number of genes selected: 

```{r}
topic <- 1
lfsr_tp <- de$lfsr[,topic]
z_tp <- de$z[,topic]
lfc_tp <- de$est[,topic]
    
index <- which(lfsr_tp < 0.05 & z_tp > 5, lfc_tp >1)
genes_pass <- rownames(de$lfsr)[index]
print(length(genes_pass))
```

There is no pathway enriched in this gene list

## k=11

### Differential analysis

The topic 2 is likely to be the unique topic for myofib

```{r}
load("/project/xinhe/xsun/heart_project/1.fast_topic/results/fast_de_modified_k11.0cellremoved.1krm.rdata")
load("/project/xinhe/xsun/heart_project/1.fast_topic/plots/volcano_plot_k11_topic2.rdata")
print(volcano_plot)
```

### Pathway enrichment analysis

We select genes of interests using a relaxed threshold: 

lfsr_tp < 0.05 & z_tp > 5, lfc_tp >1

The number of genes selected: 

```{r}
topic <- 2
lfsr_tp <- de$lfsr[,topic]
z_tp <- de$z[,topic]
lfc_tp <- de$est[,topic]
    
index <- which(lfsr_tp < 0.05 & z_tp > 5, lfc_tp >1)
genes_pass <- rownames(de$lfsr)[index]
print(length(genes_pass))
```


There is no pathway enriched in this gene list
