---
title: "Running fast topics"
author: "XSun"
date: "2024-05-15"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

# Introduction

We ran [fast topics](https://stephenslab.github.io/fastTopics/index.html) on fibroblast ATAC-seq 

```{r}
library(gridExtra)
library(Matrix)
```

# The Count Matrix  

```{r visible-code, echo=T}
load("/project/xinhe/xsun/heart_project/data/fib.acat.count.rdata")
print("dim(count)")
dim(counts)
print("counts[1:4,1:4]")
counts[1:4,1:4]
print("mean(counts > 0)")
mean(counts > 0)

row_sums <- rowSums(counts)
zero_rows <- which(row_sums == 0)
sprintf("the number of cells having 0 for all peaks = %s", length(zero_rows))

col_sums <- colSums(counts)
zero_cols <- which(col_sums == 0)
sprintf("the number of peaks having 0 for all cells = %s", length(zero_cols))
```

# Results -- cells with 0 peaks removed

We varied the number of topics from 4 to 16.

## Visualization

```{r message=FALSE, fig.width=30, fig.height=30}
load("/project/xinhe/xsun/heart_project/1.fast_topic/plots/visualization.0cellremoved.rdata")
grid.arrange(grobs = p, ncol = 4)
```

## Checking the liklihood

```{r message=FALSE, fig.width=25, fig.height=25}
load("/project/xinhe/xsun/heart_project/1.fast_topic/plots/liklihood.0cellremoved.rdata")
grid.arrange(grobs = p, ncol = 4)
```

## likelihood for each cell

```{r message=FALSE, fig.width=25, fig.height=25}
load("/project/xinhe/xsun/heart_project/1.fast_topic/plots/liklihood_dist.0cellremoved.rdata")
grid.arrange(grobs = p, ncol = 4)
```



# Results -- cells with 0 peaks removed, counts binarized

We varied the number of topics from 4 to 16.

## Visualization

```{r message=FALSE, fig.width=30, fig.height=30}
load("/project/xinhe/xsun/heart_project/1.fast_topic/plots/visualization.0cellremoved.bin.rdata")
grid.arrange(grobs = p, ncol = 4)
```

## Checking the liklihood

```{r message=FALSE, fig.width=25, fig.height=25}
load("/project/xinhe/xsun/heart_project/1.fast_topic/plots/liklihood.0cellremoved.bin.rdata")
grid.arrange(grobs = p, ncol = 4)
```

## likelihood for each cell

```{r message=FALSE, fig.width=25, fig.height=25}
load("/project/xinhe/xsun/heart_project/1.fast_topic/plots/liklihood_dist.0cellremoved.rdata")
grid.arrange(grobs = p, ncol = 4)
```



# Results -- using raw counts

We varied the number of topics from 4 to 16.

## Visualization

```{r message=FALSE, fig.width=30, fig.height=30}
load("/project/xinhe/xsun/heart_project/1.fast_topic/plots/visualization.rdata")
grid.arrange(grobs = p, ncol = 4)
```

## Checking the liklihood

```{r message=FALSE, fig.width=25, fig.height=25}
load("/project/xinhe/xsun/heart_project/1.fast_topic/plots/liklihood.rdata")
grid.arrange(grobs = p, ncol = 4)
```

## likelihood for each cell

```{r message=FALSE, fig.width=25, fig.height=25}
load("/project/xinhe/xsun/heart_project/1.fast_topic/plots/liklihood_dist.rdata")
grid.arrange(grobs = p, ncol = 4)
```